---
title: "agalluppi Homework Assignment 1"
output: html_document
---

## Section 4.1.1

### Exercise 2
####(a)
```{r}
#Create an empty vector for the first 30 values of the Fibonacci sequence and set the first two values
fib <- vector(mode = "numeric", length = 30)
fib[1] = fib[2] <- 1

#Compute values of the Fibonacci vector
for (i in 3:30){
  fib[i] <- fib[i-1] + fib[i-2]  
}

#Create an empty vector for the ratios of the Fibonacci numbers and the previous Fibonacci number
fibRatios <- vector(mode = "numeric", length = 29)

#Compute the values of the Fibonacci ratios
for (i in 2:30) {
  fibRatios[i-1] <- fib[i]/fib[i-1]
}
print(fibRatios)
```
The values do appear to be converging.


####(b)
$$ (1+\sqrt5)/2 $$

```{r, echo = FALSE}
(1+sqrt(5))/2

```
The fibRatios vector takes this value from element 16 onward.


### Exercise 3
####(a)

```{r}
answer <- 0
for (j in 1:5) answer <- answer + j
print(answer)

```

####(b)

```{r}
answer <- NULL
for (j in 1:5) answer <- c(answer, j)
print(answer)

```

####(c)

```{r}
answer <- 0
for (j in 1:5) answer <- c(answer, j)
print(answer)

```

####(d)

```{r}
answer <- 1
for (j in 1:5) answer <- answer * j
print(answer)

```

####(e)

```{r}
answer <- 3
for (j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
print(answer)

```
Knowing nothing at all about the rule, I wouldn't have confidence guessing the next terms in the sequence; however, I was able to guess that the sequence repeated given the code above.

## Section 4.1.2

### Exercise 4
```{r}
#Given initial investment, P, and number of years, y, this function caluclates interest accrued, I
interest <- function(P, y) {
   #Interest rate is 4%/yr if investment period is less than or equal to 3 years
   if (y <= 3) {
      i <- 0.04
      I <- P*((1+i)^y-1) 
   }  
  #Interest rate is 5%/yr if investment period is more than 3 years
    else {
      i <- 0.05
      I <- P*((1+i)^y-1) 
    }
  return(round(I, digits = 2))
}
```  
### Exercise 5
```{r}
#Given original principle, P, term length, m, and whether the term is open (open = TRUE, defaults to closed otherwise), this function outputs monthly mortage payment, R
MonthlyPayment <- function(P, m, open = FALSE) {
   #if the term is open, interest rate is 0.05%/month
   if (open == TRUE) {
     i <- 0.005
      R <- P*i/(1-(1+i)^(-m)) 
   }  
  #if the term is closed, interest rate is 0.04%/month
    else {
      i <- 0.004
      R <- P*i/(1-(1+i)^(-m)) 
    }
  return(round(R, digits = 2))
}
```  

## Section 4.1.3

### Exercise 2
```{r}
#Define beginning of Fibonacci sequence
Fibonacci <- c(1,1)
#While the sum of the last two elements of the Fibonacci vector is less than 300, append this sum to the end of the existing Fibonacci vector 
while(Fibonacci[length(Fibonacci)] + Fibonacci[length(Fibonacci)-1] < 300) {
  Fibonacci <- c(Fibonacci, Fibonacci[length(Fibonacci)]+Fibonacci[length(Fibonacci)-1])
}
print(Fibonacci)
  
```

### Exercise 4
```{r}
#Define starting interest rate, i, and a placeholder variable for the previous calculated value of interest, i_ref
i = 0.006
i_ref = 1
#While the last calculated interest rate, i_ref, is larger than the current value of i by more than 0.000001, assign the current value of i as i_ref and then calulate a new value for i from the current value.
while (i_ref-i > 0.000001) {
  i_ref <- i
  i <- (1-(1+i)^(-20))/19
  
}
print(i)

```

### Exercise 5
```{r}
#Define starting interest rate, i, a placeholder variable for the previous calculated value of interest, i_ref, and a variable to count the number of iterations of the while loop
i = 0.006
i_ref = 1
Iterations <- 0
#While the last calculated interest rate, i_ref, is larger than the current value of i by more than 0.000001, assign the current value of i as i_ref and then calulate a new value for i from the current value.
while (i_ref-i > 0.000001) {
  i_ref <- i
  i <- (1-(1+i)^(-20))/19
  #Adds one to the count of iterations
  Iterations <- Iterations + 1
  
}
print(i)
print(paste("Iterated", Iterations, "times."))

```

## Section 4.1.5

### Exercise 2

```{r}
Eratosthenes_sieve <- function(n) {
  # Print prime numbers up to n (based on the sieve of Eratosthenes)
  if (n >= 2) {
    sieve <- seq(2,n)
    primes <- c()
    while (length(sieve) > 0) {
      p <- sieve[1]
      primes <- c(primes, p)
      sieve <- sieve[(sieve %% p) != 0]
      if (p >= sqrt(n)) {
        break
      }
    }
    return(sieve)
  } else {
    stop("Input value of n should be at least 2.")
  }
}
```

####(c)

```{r}
Eratosthenes <- function(n) {
  # Print prime numbers up to n (based on the sieve of Eratosthenes)
  if (n >= 2) {
    sieve <- seq(2,n)
    primes <- c()
    while (length(sieve) > 0) {
      p <- sieve[1]
      primes <- c(primes, p)
      sieve <- sieve[(sieve %% p) != 0]
      if (p >= sqrt(n)) {
        break
      }
    }
    return(c(primes,sieve))
  } else {
    stop("Input value of n should be at least 2.")
  }
}

```


