---
title: "agalluppi Homework Assignment 1"
output: html_document
---

## Section 4.1.1

### Exercise 2
####(a)
```{r}
#Create an empty vector for the first 30 values of the Fibonacci sequence and set the first two values
fib <- vector(mode = "numeric", length = 30)
fib[1] = fib[2] <- 1

#Compute values of the Fibonacci vector
for (i in 3:30){
  fib[i] <- fib[i-1] + fib[i-2]  
}

#Create an empty vector for the ratios of the Fibonacci numbers and the previous Fibonacci number
fibRatios <- vector(mode = "numeric", length = 29)

#Compute the values of the Fibonacci ratios
for (i in 2:30) {
  fibRatios[i-1] <- fib[i]/fib[i-1]
}
print(fibRatios)
```
The values do appear to be converging.


####(b)
$$ (1+\sqrt5)/2 $$

```{r, echo = FALSE}
(1+sqrt(5))/2

```
The fibRatios vector takes this value from element 16 onward.


### Exercise 3
####(a)

```{r}
answer <- 0
for (j in 1:5) answer <- answer + j
print(answer)

```

####(b)

```{r}
answer <- NULL
for (j in 1:5) answer <- c(answer, j)
print(answer)

```

####(c)

```{r}
answer <- 0
for (j in 1:5) answer <- c(answer, j)
print(answer)

```

####(d)

```{r}
answer <- 1
for (j in 1:5) answer <- answer * j
print(answer)

```

####(e)

```{r}
answer <- 3
for (j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
print(answer)

```
Knowing nothing at all about the rule, I wouldn't have confidence guessing the next terms in the sequence; however, I was able to guess that the sequence repeated given the code above.

## Section 4.1.2

### Exercise 4
```{r}
#Given initial investment, P, and number of years, y, this function caluclates interest accrued, I
interest <- function(P, y) {
   #Interest rate is 4%/yr if investment period is less than or equal to 3 years
   if (y <= 3) {
      i <- 0.04
      I <- P*((1+i)^y-1) 
   }  
  #Interest rate is 5%/yr if investment period is more than 3 years
    else {
      i <- 0.05
      I <- P*((1+i)^y-1) 
    }
  return(round(I, digits = 2))
}
```  
### Exercise 5
```{r}
#Given original principle, P, term length, m, and whether the term is open (open = TRUE, defaults to closed otherwise), this function outputs monthly mortage payment, R
MonthlyPayment <- function(P, m, open = FALSE) {
   #if the term is open, interest rate is 0.05%/month
   if (open == TRUE) {
     i <- 0.005
      R <- P*i/(1-(1+i)^(-m)) 
   }  
  #if the term is closed, interest rate is 0.04%/month
    else {
      i <- 0.004
      R <- P*i/(1-(1+i)^(-m)) 
    }
  return(round(R, digits = 2))
}
```  

## Section 4.1.3

### Exercise 2
```{r}
#Define beginning of Fibonacci sequence
Fibonacci <- c(1,1)
#While the sum of the last two elements of the Fibonacci vector is less than 300, append this sum to the end of the existing Fibonacci vector 
while(Fibonacci[length(Fibonacci)] + Fibonacci[length(Fibonacci)-1] < 300) {
  Fibonacci <- c(Fibonacci, Fibonacci[length(Fibonacci)]+Fibonacci[length(Fibonacci)-1])
}
print(Fibonacci)
  
```

### Exercise 4
```{r}
#Define starting interest rate, i, and a placeholder variable for the previous calculated value of interest, i_ref
i = 0.006
i_ref = 1
#While the last calculated interest rate, i_ref, is larger than the current value of i by more than 0.000001, assign the current value of i as i_ref and then calulate a new value for i from the current value.
while (i_ref-i > 0.000001) {
  i_ref <- i
  i <- (1-(1+i)^(-20))/19
  
}
print(i)

```

### Exercise 5
```{r}
#Define starting interest rate, i, a placeholder variable for the previous calculated value of interest, i_ref, and a variable to count the number of iterations of the while loop
i = 0.006
i_ref = 1
Iterations <- 0
#While the last calculated interest rate, i_ref, is larger than the current value of i by more than 0.000001, assign the current value of i as i_ref and then calulate a new value for i from the current value.
while (i_ref-i > 0.000001) {
  i_ref <- i
  i <- (1-(1+i)^(-20))/19
  #Adds one to the count of iterations
  Iterations <- Iterations + 1
  
}
print(i)
print(paste("Iterated", Iterations, "times."))

```

## Section 4.1.5

### Exercise 2

####(a)
The function assigns a "sieve" vector from 2 to n, then takes the first element in sieve and adds it to the "primes" vector.  It then removes all multiples of the first element in sieve from sieve, and repeats until there are no more numbers in sieve.  The primes vector is now all the prime numbers originally in sieve.

####(b)
```{r}
Eratosthenes_sieve <- function(n) {
  # Print sieve when p >= sqrt(n)
  if (n >= 2) {
    sieve <- seq(2,n)
    primes <- c()
    while (length(sieve) > 0) {
      p <- sieve[1]
      primes <- c(primes, p)
      sieve <- sieve[(sieve %% p) != 0]
      # Added a break when p >= sqrt(n) to show that sieve is all prime at this point
      if (p >= sqrt(n)) {
        break
      }
    }
    return(sieve)
  } else {
    stop("Input value of n should be at least 2.")
  }
}
```

####(c)

```{r}
Eratosthenes <- function(n) {
  # Print prime numbers up to n (based on the sieve of Eratosthenes) using the more efficient "if" statement break above
  if (n >= 2) {
    sieve <- seq(2,n)
    primes <- c()
    while (length(sieve) > 0) {
      p <- sieve[1]
      primes <- c(primes, p)
      sieve <- sieve[(sieve %% p) != 0]
      if (p >= sqrt(n)) {
        break
      }
    }
    return(c(primes,sieve))
  } else {
    stop("Input value of n should be at least 2.")
  }
}

```

## Section 4.2.1

### Exercise 2

####(a)
```{r}
# This function takes initial deposit, P, interest rate, i.r, and number of months, n, and calculated the total amount in the account with interest compounded after n months.
compound.interest <- function(P, i.r, n) {
  for (i in n) {
    P <- P*(1+i.r)^n 
  }
  return(P)
}

```

####(b)
```{r, echo = FALSE}
# This function takes initial deposit, P, interest rate, i.r, and number of months, n, and calculated the total amount in the account with interest compounded after n months.
compound.interest <- function(P, i.r, n) {
  for (i in n) {
    P <- P*(1+i.r)^n 
  }
  return(P)
}
compound.interest(1000, 0.01, 30)
```

### Exercise 3
```{r}
# Define the function for which you want to find zeroes
f_x <- function(x)
  x^2 - 1

# This function uses the bisection algorithm to find zeroes of the function f_x given two inital boundaries a and b.
bisection <- function(a, b) {
  repeat {
  # Find the midpoint of the boundaries, c, and calculate the value of the function at a,b, and c
  c <- (a+b)/2
  f_c <- f_x(c)
  f_a <- f_x(a)
  f_b <- f_x(b)
  
  # Stops the function if a and b do not enclose a zero or enclose more than one zero
  if (f_a < 0 && f_b < 0 || f_a > 0 && f_b > 0) {
    stop("a and b either do not enclose a zero or enclose more than one zero")
  }
  
  # When f(c) if sufficiently close to zero, stop iterating and return c 
    if (f_c == 0 || abs(f_c) < 0.000001) {
      break
    }
  # Otherwise, perform the appropriate replacement of an endpoint per the bisection algorithm and continue
    else if (f_a > 0 && f_c < 0 || f_a < 0 && f_c > 0) {
    b <- c
    f_b <- f_c
    
  }
    else if (f_b > 0 && f_c < 0 || f_b < 0 && f_c > 0) {
    a <- c
    f_a <- f_c
  }
  }
  return(c)
}
```

## Section 4.4.1

### Exercise 1
```{r}
# 1. use a merge sort to sort a vector, x. If decreasing is "TRUE," the vector will be sorted in descending order; otherwise, it will be sorted in ascending order.
mergesort <- function(x, decreasing = FALSE) {
  # Check for a vector that doesn't need sorting
  len <- length(x)
  if (len < 2) result <- x
  else {
    if (decreasing == "TRUE") {
    # 2. sort x into result
    # 2.1. split x in half
    y <- x[1:(len %/% 2)]
    z <- x[(len %/% 2 + 1):len]
    # 2.2. sort y and z
    y <- mergesort(y)
    z <- mergesort(z)
    # 2.3 merge y and z into a sorted result
    result <- c()
    # 2.3.1. while (some are left in both piles)
    while (min(length(y), length(z)) > 0) {
      # 2.3.2 put the largest first element on the end
      # 2.3.3. remove it from y or z
      if (y[length(y)] > z[length(z)]) {
        result <- c(result, y[length(y)])
        y <- y[length(y)-1]
      } else {
          result <- c(result, z[length(z)])
          z <- z[length(z)-1]
      }
    }
      # 2.3.4. put the leftovers onto the end of result
  if (length(y) > 0)
        result <- c(result,y)
  else 
        result <- c(result, z)
    } else {
    # 2. sort x into result
    # 2.1. split x in half
    y <- x[1:(len %/% 2)]
    z <- x[(len %/% 2 + 1):len]
    # 2.2. sort y and z
    y <- mergesort(y)
    z <- mergesort(z)
    # 2.3 merge y and z into a sorted result
    result <- c()
    # 2.3.1. while (some are left in both piles)
    while (min(length(y), length(z)) > 0) {
      # 2.3.2 put the smallest first element on the end
      # 2.3.3. remove it from y or z
      if (y[1] < z[1]) {
        result <- c(result, y[1])
        y <- y[-1]
      } else {
          result <- c(result, z[1])
          z <- z[-1]
      }
    }
      # 2.3.4. put the leftovers onto the end of result
  if (length(y) > 0)
        result <- c(result,y)
  else 
        result <- c(result, z)
    }
      
  }
      return(result)
}

```

### Exercise 2

```{r}
fx_n <- fy_n <- 1

gx_n = function(x,y){
  2*x + 2*y^2
}
gy_n = function(x,y) {
  x^2 + 4*y
}
f_n = function(x,y) {
  x+y
}
g_n = function(x,y) {
  x^2 + 2*y^2 - 2
}
d_n = function(x,y) {
  fx_n*gy_n(x,y) - fy_n*gx_n(x,y)
}

Newton <- function(x,y) {
while (abs(x+y) > 0.000001 || abs(x^2+2*y^2-2) > 0.000001) {
  x <- x - (gy_n(x,y)*f_n(x,y) - fy_n*g_n(x,y))/d_n(x,y)
  y <- y - (fx_n*g_n(x,y) - gx_n(x,y)*f_n(x,y))/d_n(x,y)
}
return(c(x,y))
}

```


## Chapter 4 Exercises

### Exercise 1
```{r}
# Given the value of x and a vector of constants, c, this function evaluates a polnomial of order length(c)-1
directpoly <- function(x, c) {
 P <- 0
  for (i in 1:length(c)) {
    P <- P + c[i]*x^(length(c)-i)
  }
  return(P)
}
```

### Exercise 2
```{r}
# This function uses Horner's rule to exaluate a polynomial expression given the value of x and a vector of its constants, c.
hornerpoly <- function(x, c) {
  a <- c
  for (i in 2:length(c)) {
    a[i] <- a[i-1]*x+c[i]
  }
  return(a[length(c)])
}

```

### Exercise 3

####(a)
```{r, echo = FALSE}
print(paste("Directpoly time"))
system.time(directpoly(x=seq(-10, 10, length=5000000),
    c(1, -2, 2, 3, 4, 6, 7)))
print(paste("Hornerpoly time"))
system.time(hornerpoly(x=seq(-10, 10, length=5000000),
    c(1, -2, 2, 3, 4, 6, 7)))

```

####(b)
```{r, echo = FALSE}
print(paste("Directpoly time"))
system.time(directpoly(x=seq(-10, 10, length=5000000),
    c(2, 17, -3)))
print(paste("Hornerpoly time"))
system.time(hornerpoly(x=seq(-10, 10, length=5000000),
    c(2, 17, -3)))

```